---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

# defraEmissionsFactors

<!-- badges: start -->
<!-- badges: end -->

The goal of defraEmissionsFactors is to make emissions factors provided by DEFRA more easily accessible and transparent.

The starting point is [Greenhouse gas reporting: conversion factors 2018](https://www.gov.uk/government/publications/greenhouse-gas-reporting-conversion-factors-2018).
The following code chunk downloads and extracts relevant data, in this case kg CO2 per km, for different car sizes:

```{r cars, message=FALSE}
library(tidyverse)
u2018 = "https://www.gov.uk/government/uploads/system/uploads/attachment_data/file/715425/Conversion_Factors_2018_-_Condensed_set__for_most_users__v01-01.xls"
f = file.path(tempdir(), "Conversion_Factors_2018_-_Condensed_set__for_most_users__v01-01.xls")
download.file(u2018, f)
cars_size = readxl::read_excel(f, sheet = "Passenger vehicles", skip = 45) %>%
  rename_all(snakecase::to_snake_case) %>% 
  rename_all(gsub, pattern = "kg_|_[0-9]+$", replacement = "")
propulsion_types = c("petrol", "diesel", "hybrid", "CNG", "LPG", "Unknown", "phev", "bev")
names(cars_size)[5:36] = paste(names(cars_size[5:36]), rep(propulsion_types, each = 4), sep = "_")
cars_size_co2 = cars_size %>% 
  select(-(1:2)) %>%
  select(matches("type|unit|co_2_[d|p|h]")) %>% 
  filter(unit == "km") %>% 
  slice(1:4) %>% 
  select(-unit) %>% 
    mutate(co_2_petrol = as.numeric(co_2_petrol))
knitr::kable(cars_size_co2, digits = 3)
```

This shows that cars typically emit around 180 g CO2/km, nearly double the EU's [2021 target](https://ec.europa.eu/clima/policies/transport/vehicles/cars_en) of 95 g CO2/km.
We can convert these into energy use values based on the following conversion factors:

Burning diesel emits 73.6 gCO2/MJ ([source](https://enpos.weebly.com/uploads/3/6/7/2/3672459/co2_direct_combustion_jokiniemi.pdf)), meaning that 1 KG of CO2 released from burning diesel is associated with `1 / 0.0726` (13.6) MJ of energy use.
This figure, and the associated value of `0.0728` for petrol, means we can calculate energy use of different types of cars as follows:

```{r}
cars_energy = cars_size_co2 %>% 
  mutate(energy_diesel = co_2_diesel / 0.0726) %>% 
  mutate(energy_petrol = co_2_petrol / 0.0728) %>% 
  select(type, matches("energy"))
```

Crudely assuming an even petrol/diesel fuel mix, we can estimate the average energy use per km driven as follows:

```{r}
(cars_energy$energy_diesel[4] + cars_energy$energy_petrol[4]) / 2
```


```{r, eval=FALSE, echo=FALSE}
# removed as gets the wrong values
fuel_data_kwh = readxl::read_excel(f, sheet = "Fuels") %>% 
  fill(`...3`, .direction = "up") %>% 
  filter(`...4` == "kWh (Gross CV)") %>% 
  select(3:6) %>% 
  filter(grepl(pattern = "Diesel|Petrol", `...3`)) %>% 
  mutate_at(vars(3:4), as.numeric)
names(fuel_data_kwh) = c("fuel", "energy_unit", "kg_co2e", "kg_co2")
knitr::kable(fuel_data_kwh)

conversion_kwh_petrol = fuel_data_kwh %>% 
  filter(fuel == "Petrol (100% mineral petrol)") %>% 
  pull(kg_co2)
conversion_kwh_diesel = fuel_data_kwh %>% 
  filter(fuel == "Diesel (100% mineral diesel)") %>% 
  pull(kg_co2)
cars_energy_kwh = cars_size_co2 %>% 
  select(type, co_2_petrol, co_2_diesel) %>% 
  mutate(energy_petrol = co_2_petrol / conversion_kwh_petrol) %>% 
  mutate(energy_diesel = co_2_diesel / conversion_kwh_diesel) %>% 
  select(matches("type|energy")) 
knitr::kable(cars_energy_kwh, digits = 3)
# And the final step is to convert these values into SI units for energy analysis:
cars_energy = cars_energy_kwh %>% 
  mutate_if(is.numeric, function(x) x / 3.6)
  
```

